Adopting agile software development in the large organizations: A systematic literature review

### 2014-06-07 journal/article.tex
Threats to validity; research process figures

The research process is presented starting from the research questions. After
that we present the definition of large scale which this research builds upon.
The bulk of this section concentrates on presenting the steps of the research
process, starting from the definition of inclusion criteria, continuing with
describing how the primary studies were selected, and finally presenting the
method for synthesizing the primary studies. The section concludes with a
discussion on limitations of the method.

@@

The research was conducted as an application of Kitchenham's
\cite{Kitchenham2007} guidelines for systematic literature review. Selection of
primary studies was done as a keyword based database search and manual filtering
of included studies. The filtering process was executed independently by two
researchers (Dikert, Paasivaara). The selected primary studies were coded for
qualitative data extraction and analysis. The entire process was audited and
mentored by a third researcher (Lassenius).

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=1\textwidth]{graphics/research_process_full.pdf}
    \caption{Full study selection process}
    \label{fig:selectionprocess_full}
  \end{center}
\end{figure}

TODO:
* Duplicates removed?
* Ei voi ymmärtää miten preliminary search filtteröitiin

@@

\subsubsection{Coding of primary documents}
\label{sec:coding}

The primary documents were coded using the Atlas TI qualitative data analysis software.
An integrated deductive and inductive approach was chosen for coding the primary
studies \cite{Cruzes2011a}. The coding was designed to have a contextual part
and a findings part as also suggested by Cruzes and Dybå \cite{Cruzes2011a}. A
list of codes was established for contextual information, as presented in Table
\ref{table:contextualcodes}. Codes related to the research questions were chosen
to be created by an inductive process. The reason for inducting codes rather
than using an a priori approach was to avoid the researchers' previous
assumptions of the research area to affect the choice of codes

@@

\begin{table}[h]
    \centering
    \begin{tabular}{ l r r }
        \toprule
        Code family    &  Codes  &  Quotations
        \\
        \midrule
        RQ1: Reason to change &        30 &  123 \\
		RQ2: Transformation process &  16 &  580 \\
		RQ3: Challenges &              40 &  323 \\
		RQ3: Success factors &         44 &  260 \\
		Investing in change  &          5 &  137 \\
		Practices &                    11 &  170 \\
		Contextual &                    8 &  215 \\
		Total &                       154 & 1575 \\
        \bottomrule
    \end{tabular}
    \caption{Number of codes and quotations per code family}
    \label{table:codecount}
\end{table}

@@

Coding was done by only one researcher due to resource constraints and volume of
source material. As recommended by Kitchenham \cite{Kitchenham2007} we conducted
an experiment with the goal to assess the consistency of coding that two
independent researchers may achieve.
In the experiment Dikert and Paasivaara coded independently a sample of five
primary documents. The sample was hand picked as a representative subset of the
source material. The coding was done relying solely on the guidelines presented
in the previous chapter, and the researchers were careful not to discuss the
choice of codes or interpretations relating to codes before the experiment was
completed. Finally the independent codings of the two researchers were analyzed
with two metrics: by the number of matching themes found, and by the amount of
overlap in quotations similar codes have.
% See Kitcenham, chapter 6.4

@@

\subsection{Limitations of research method}

-- A keyword search can not identify the facet of \emph{large scale}, and also
   \emph{transformation} is difficult to capture with keywords. Because of this
   we assume that it is likely that a some studies have been missed by the
   keyword search. As an improvement we would suggest combining the entire
   archive of \emph{Agile Conference} and \emph{Agile Processes in Software
   Engineering and Extreme Programming} to the results of the keyword search.
   These two conferences provided more than two thirds of the primary sources.

-- Few of the primary studies presented a research method. Most of the studies
   were experience reports where the author was a member of the organization
   discussed. Because of this we must assume that the primary studies are
   biased. Due to this risk of bias we chose not to make statistical
   interpretations of the results. We only use relative measure when comparing
   results.

-- We avoid numbers in results, such as frequency of a phenomena, except in the
   meta-analysis section.

Bias  --  systematic error

Internal quality  --  validity

External quality  --  generalizability

>>>>

The research process is presented starting from the research questions. After
that we present the definition of large scale which this research builds upon.
The bulk of this section concentrates on presenting the steps of the research
process, starting from the definition of inclusion criteria, continuing with
describing how the primary studies were selected, and finally presenting the
method for analyzing and synthesizing the primary studies. The section concludes
with a discussion on limitations of the method.

@@

The research was conducted as an application of Kitchenham's
\cite{Kitchenham2007} guidelines for systematic literature review. Selection of
primary studies was done using first a keyword based database search to identify
potentially relevant sources, and then manually filtering the search result. The
filtering process was executed independently by two researchers (Dikert,
Paasivaara). The selected primary studies were coded for qualitative data
extraction by one researcher (Dikert). The results of the study were finally
elicitated by aggregating and analyzing the coding of the primary documents.
Figure \ref{fig:researchprocess} presents an outline of the research process.
The entire process was audited and mentored by a third researcher (Lassenius).

\begin{figure}[b]
  \begin{center}
    \includegraphics{graphics/research_process.pdf}
    \caption{Outline of the research process}
    \label{fig:researchprocess}
  \end{center}
\end{figure}

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=1\textwidth]{graphics/research_process_full.pdf}
    \caption{The study selection process}
    \label{fig:selectionprocess_full}
  \end{center}
\end{figure}

@@

\subsubsection{Coding of primary studies}
\label{sec:coding}

The primary studies were coded using the Atlas TI qualitative data analysis software.
An integrated deductive and inductive approach was chosen for coding the primary
studies \cite{Cruzes2011a}. The coding was designed to have a contextual part
and a findings part as also suggested by Cruzes and Dybå \cite{Cruzes2011a}. A
list of codes was established for contextual information, as presented in Table
\ref{table:contextualcodes}. Codes related to the research questions were chosen
to be created by an inductive process. The reason for inducting codes rather
than using an a priori approach was to avoid the researchers' previous
assumptions of the research area to affect the choice of codes

@@

\begin{table}[t]
    \centering
    \begin{tabular}{ l r r }
        \toprule
        Code family    &  Codes  &  Quotations
        \\
        \midrule
        RQ1: Reason to change &        30 &  123 \\
		RQ2: Transformation process &  16 &  580 \\
		RQ3: Challenges &              40 &  323 \\
		RQ3: Success factors &         44 &  260 \\
		Investing in change  &          5 &  137 \\
		Practices &                    11 &  170 \\
		Contextual &                    8 &  215 \\
		Total &                       154 & 1575 \\
        \bottomrule
    \end{tabular}
    \caption{Number of codes and quotations per code family}
    \label{table:codecount}
\end{table}

@@

Coding was done by only one researcher due to resource constraints and volume of
source material. As recommended by Kitchenham \cite{Kitchenham2007} we conducted
an experiment with the goal to assess the consistency of coding that two
independent researchers may achieve.
In the experiment Dikert and Paasivaara coded independently a sample of five
primary studies. The sample was hand picked as a representative subset of the
source material. The coding was done relying solely on the guidelines presented
in the previous chapter, and the researchers were careful not to discuss the
choice of codes or interpretations relating to codes before the experiment was
completed. Finally the independent codings of the two researchers were analyzed
with two metrics: by the number of matching themes found, and by the amount of
overlap in quotations similar codes have.
% See Kitcenham, chapter 6.4

@@

\subsection{Threats to validity and limitations}

To assess the validity of this research internal validity and external validity
needs to be considered. In this section we present the identified threats to
validity and the strategies used to mitigate them. We highlight two specific
limitations in the research: weakness in keyword search, and the strength of
evidence of primary sources.

The threat to internal validity is subjective bias, which affects two parts of
the research. First, the selection of the primary studies may have been
distorted by interpreting the inclusion criteria falsely. This risk was
mitigated at start by using three researchers in designing the inclusion
criteria. When the inclusion criteria was subsequently applied the initial
filtering was performed independently by two researchers, and unclear cases were
resolved by case-by-case discussions between two or three researchers.

The second part of the research that may have been affected by subjective bias
was the elicitation of results by coding and analysis. Our tools for mitigating
this threat were limited, as the work stages in question are particularly
laborous and for resource constraints the process could not be duplicated. We
assessed the bias of the coding step using an experiment, and prevented
subjective bias in analysis by making the results as tracable as possible. The
coding experiment revealed that two researchers can find roughly 2 of 3 themes
in an experience report identically. The traceability of results is achieved by
supplying references to each claim presented in the results.

This study had only a minor threat to external validity. The systematic
literature review approach includes steps to identify all studies that may be
relevant in the study area. As long as the systematic approach is adhered to
threats against external validity are minimized. External validity was also
confirmed by saturation of the primary studies. Saturation was evident, as when
the majority of the primary studies had been coded almost all quotations for the
remaining studies could be assigned to existing codes.

A particular problem in this systematic literature review was the limitations of
boolean keyword searches in online databases. A keyword search can not identify
the facet of \emph{large scale}, and also \emph{transformation} is difficult to
capture with keywords. Because of this we assume that it is likely that a some
studies have been missed by the keyword search, potentially causing bias. To
eliminate the potential bias we would suggest combining the entire archive of
\emph{Agile Conference} and \emph{Agile Processes in Software Engineering and
Extreme Programming} to the results of the keyword search. These two conferences
provided more than two thirds of the primary sources.

Another particular problem was that few of the primary studies presented a
research method. Most of the studies were experience reports where the author
was a member of the organization discussed. Because of this we must assume that
the primary studies are biased to some extent. Due to this risk of bias we chose
not to make quantitative interpretations in the results, and use only
qualitative analysis. We conclude that in the current state of research it is
necessary to include studies with varying strength of evidence in order to
relevantly aggregate evidence on large scale transformations.
